Glyph-maps for Visually Exploring Temporal Patterns in Climate Data and Models

### 2012-03-05 revision/paper.tex
changed figure 5 caption
Model desc tweaks

Alternatively, local scale can be used. In local scaling, the values within each location are scaled to range $[0, 1]$ before plot construction. This makes it easier to compare shape (ignoring amplitude), and draws attention to locations with large relative variation. Figure~\ref{fig:scaling} compares global scaling and local scaling for the smoothed temperature data from the EXPO data. (The model used to get the temperature values is a more complicated than that used previously to de-seasonalize. A generalized additive model using a month, and a smoothed day, term \citep{wood:2006} was fit to the observed data. The smoothed day was computed from the monthly resolution. Predicted values were computed for each day of the six year period. Season is removed but not overall average and long-term and trend.) On the global scale, the El Ni\~no blip in temperature is just visible, along with variation in the average temperature at each location. Local scaling emphasizes the individual shapes: the impact of El Ni\~no can be seen to be over a wider region, linear increases across the Andes, and decreases in the Gulf of Mexico.

@@

  \caption{Smoothed de-seasonalized daily temperature glyph-map. (Top-left)
globally scaled, (Top-right) locally scaled. (Bottom) Legend showing ranges of individual glyphs.}
  \label{fig:scaling}
\end{figure}

>>>>

Alternatively, local scale can be used. In local scaling, the values within each location are scaled to range $[0, 1]$ before plot construction. This makes it easier to compare shape (ignoring amplitude), and draws attention to locations with large relative variation. Figure~\ref{fig:scaling} compares global scaling and local scaling for the smoothed temperature data from the EXPO data. (The model is more complicated than the model used previously to de-seasonalize the data. A generalized additive model \citep{wood:2006} with qualitative month and smoothed day terms was fit to the observed data and predicted values were computed for each day of the six year period. Seasonal effects are removed but not overall average or long-term trend.) On the global scale, the El Ni\~no blip in temperature is just visible, along with variation in the average temperature at each location. Local scaling emphasizes the individual shapes: the impact of El Ni\~no can be seen to be over a wider region, linear increases across the Andes, and decreases in the Gulf of Mexico.

@@

  \caption{Glyph-map using predicted daily temperature values from a generalized additive model fit, where seasonality has been extracted. (Top-left)
globally scaled, (Top-right) locally scaled. (Bottom) Legend showing ranges of individual glyphs.}
  \label{fig:scaling}
\end{figure}
